Bulk Semantic Annotation with an Partially Known Knowledge Base

### 2024-11-10 doc/sysdesc.tex
more

\documentclass{llncs}
\usepackage[show]{ed}
\usepackage{cleveref}
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{stex-logo}

@@

\newcommand\ALeA{\textsf{ALeA}\xspace}
\newcommand\srf{\textsf{srify}\xspace}

@@

In this paper, we address tools that help attack the coverage problem in symbolic/semantic
approaches in practice. The producer task we use as a case study is that of adaptive
learning assistants for MINT subjects, concretely the \ALeA system
\cite{BerBetChu:lssmkm23}, which uses \sTeX \cite{MueKo:sdstex22,sTeX:github:on} -- a
variant of {\LaTeX} that allows to embed semantic annotations -- for a knowledge
representation and generates learner-adaptive learning objects instrumented with
learning-support interactions from that. The \textbf{\sTeX/\ALeA content commons}
\ednote{MK@MK: continue to describe the size and distribution, and authorship}

@@

The main practical problem of annotating presentational {\LaTeX} course materials into
semantic documents that carry enough information to support meaningful learning support
services is to reference

>>>>

\documentclass{llncs}
\usepackage[show]{ed}
\usepackage{cleveref}
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{stex-logo}
\usepackage{listings}
\definecolor{codegray}{rgb}{0.9,0.9,0.9}
\lstset{basicstyle=\sf,columns=fullflexible,backgroundcolor = \color{codegray}}

@@

\newcommand\ALeA{\textsf{ALeA}\xspace}
\newcommand\srf{\textsf{srify}\xspace}
\def\llangle{\langle\kern-.2em\langle}
\def\rrangle{\rangle\kern-.2em\rangle}

@@

In this paper, we address tools that help attack the coverage problem in symbolic/semantic
approaches in practice. The producer task we use as a case study is that of adaptive
learning assistants for MINT subjects, concretely the \ALeA system
\cite{BerBetChu:lssmkm23}, which uses \sTeX \cite{MueKo:sdstex22,sTeX:github:on} -- a
variant of {\LaTeX} that allows to embed semantic annotations -- for a knowledge
representation and generates learner-adaptive learning objects instrumented with
learning-support interactions from that. The \textbf{\sTeX/\ALeA content commons}
\ednote{MK@MK: continue to describe the size and distribution, and
  authorship}\ednote{explain the \textbf{domain model} (the SMGloM) and the
  \textbf{formulation model} (cf. \cite{BerBetChu:lssmkm23}) in the commons and give their
  releative sizes.}

@@

The main practical problem of annotating presentational {\LaTeX} course materials into
semantic documents that carry enough information to support meaningful learning support
services is to add semantic references to technical terms from the domain jargon. Take for
instance a learning object like the paragraph in \cref{fig:lo}, where we want to annotate
the the term ``state space'' with the (semantic) concept of a ``state space of a search
problem''. The latter can either be defined earlier in the course or in the domain mode
(or both).

\begin{figure}[ht]\centering
  \includegraphics[width=8cm]{problem}
  \caption{An annotated Learning Object from an AI Lecture}\label{fig:lo}
\end{figure}

Concretely this is about converting the {\LaTeX} string
\begin{lstlisting}[language={[LaTeX]TeX}]
\item \blue{Problem:} The state space has local minima, e.g. the board on the
  right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
into the (partially annotated item:  
\begin{lstlisting}[language={[LaTeX]TeX},morekeywords={sn,importmodule}]
\importmodule[smglom/search]{mod?state-space}
[...]  
\item \blue{Problem:} The \sn{state space} has local minima, e.g. the
  board on the right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
Concretely, the author has to remember the module for the term ``state space'' from
\cref{fig:state-space}, the symbol name \lstinline|state space|, annotate it with the
\sTeX macro \lstinline|\sn{...}|, and add the module import in the first line (unless it
is already imported). In this case the symbol name happens to be the same as its English
verbalization, otherwise we would have to use
\lstinline[mathescape]|\sr{$\llangle symbol name\rrangle$}{state space}|.

\begin{figure}[ht]\centering
  \includegraphics[width=12cm]{state-space}
  \caption{The definition of the term ``state space'' from the domain model}\label{fig:state-space}
\end{figure}

For this, the author has to be aware that the module from \cref{fig:state-space} exists,
the symbol name, know the URL and manage reduncancy of the imports. At ca 150.000 words in
a semester's words of lecture notes and ca. 10-15\% of technical terms a rather daunting
task.
